Dynamic Walking and Whole-Body Motion Planning for Humanoid Robots: an Integrated Approach

### 2012-12-05 paper.tex
Add RRT* reference
Minor changes following reviewer comments

When planning collision free motions for humanoid robots, different representations
of the robot and its environment can be used. The choice of the level of details of
the representation indicates the difficulty of the considered problem. The simplest
option consists in considering the robot as a navigating 2D shape, and computing 
obstacle avoidance in a planar model of the world. Another possibility is to 
compute only collision-free footsteps. 
In complex and difficult environments, such as the one presented in Fig.~\ref{fig:couv},
it can be necessary to consider exact 3D models of a humanoid robot and its environment.

@@

There are two main ways of using motion planners to generate dynamically balanced robotic motions.
The more general one is to plan in a robot dynamic space. By taking into account both robot
configuration and velocity, motions that satisfy dynamic balance constraints can be generated
at a planning phase. When planning motion for humanoid robots, this is a particularly costly
approach, as the size of the space to explore is augmented with the robot velocity and footstep
positions. The other way is to first plan  a geometric path that can be approximated by a
dynamic trajectory in a second step. The approach we present in this paper falls into the second 
category. Some feasible dynamic motions are inherently impossible to compute with this kind
of approach. For example, jumping motions cannot be generated by a purely geometric planner.

@@

The problem of inverse kinematics for a humanoid robot, or any articulated
structure, is to compute a joint position to achieve an end-effector pose. As the
robots we deal with are redundant, it is natural to take advantage of
this redundancy by specifying multiple tasks, potentially with
different priorities. This problem has been widely studied in robotics
planning and control literature, and many Jacobian-based solutions have been
proposed, among which 
\cite{nakamura1986iks}, \cite{siciliano1991gfm},
\cite{baerlocher1998tpf} and \cite{khatib2004wbd}.
Obstacle avoidance can be taken into account with similar methods. To
do so, one has to include the obstacles as  constraints to
satisfy, see for example \cite{kanehiro2008lca}.
These methods are prone to fall into local minima, thus global motion
planning is needed to overcome this limitation.
\cite{TouGieGoe2007} propose a motion generation method where tasks follow
trajectories defined by cubic B-splines. The whole-body motion is optimized
with respect to the control point positions. This method can take into account
collisions with simple obstacles.

@@

Our algorithm generalizes randomized tree expansion strategies, introduced in both \cite{HsuLat99c} and
\cite{kuffner00rrtconnect} to constrained motion planning problems. Next subsection will recall the structure of the RRT algorithm,
a popular randomized motion planning algorithm.

@@

>>>>

When planning collision free motions for humanoid robots, different representations
of the robot and its environment can be used. The choice of the level of details of
the representation indicates the difficulty of the considered problem. The simplest
option consists in considering the robot as a navigating 2D shape \cite{pettre20032}, and computing 
obstacle avoidance in a planar model of the world. Another possibility is to 
compute only collision-free footsteps  \cite{kuffner2001footstep}. 
In complex and difficult environments, such as the one presented in Fig.~\ref{fig:couv},
it can be necessary to consider exact 3D models of a humanoid robot and its environment.

@@

There are two main ways of using motion planners to generate dynamically balanced robotic motions.
The more general one is to plan in a robot dynamic space, see for example \cite{shkolnik2011bounding}. 
By taking into account both robot
configuration and velocity, motions that satisfy dynamic balance constraints can be generated
at a planning phase. When planning motion for humanoid robots, this is a particularly costly
approach, as the size of the space to explore is augmented with the robot velocity and footstep
positions. The other way is to first plan  a geometric path that can be approximated by a
dynamic trajectory in a second step \cite{yoshida-humanoids05}. The approach we present in this paper 
falls into the second 
category. Some feasible dynamic motions are inherently impossible to compute with this kind
of approach. For example, jumping motions cannot be generated by a purely geometric planner.

@@

The problem of inverse kinematics for a humanoid robot, or any articulated
structure, is to compute a joint position to achieve an end-effector pose. As the
robots we deal with are redundant, it is natural to take advantage of
this redundancy by specifying multiple tasks, potentially with
different priorities. This problem has been widely studied in robotics
planning and control literature, and many Jacobian-based solutions have been
proposed, among which 
\cite{nakamura1986iks}, \cite{siciliano1991gfm},
\cite{baerlocher1998tpf} and \cite{khatib2004wbd}.
Obstacle avoidance can be taken into account with similar methods. To
do so, one has to include the obstacles as  constraints to
satisfy, see for example \cite{kanehiro2008lca}.
These methods are prone to fall into local minima, thus global motion
planning is needed to overcome this limitation. 
{\bf Note that when local methods find solutions, these are usually smoother
and may look more natural. The choice of using global motion planners
is justified by the need for complete algorithms.}
\cite{TouGieGoe2007} propose a motion generation method where tasks follow
trajectories defined by cubic B-splines. The whole-body motion is optimized
with respect to the control point positions. This method can take into account
collisions with simple obstacles.

@@

Our algorithm generalizes randomized tree expansion strategies, introduced in both \cite{HsuLat99c} and
\cite{kuffner00rrtconnect} to constrained motion planning problems. Next subsection will recall the structure of the RRT algorithm,
a popular randomized motion planning algorithm. 
{\bf Our method could be applied to any randomized sampling planning method, such
as RRT\up{*} \cite{Karaman01062011}, which is slower than RRT but generates optimal
paths.}

@@

{\bf \cite{berenson2011task} presents a formal proof that projection-based 
constrained random motion planning on a fixed dimension manifold is 
probabilisticaly complete. This proof applies to our algorithm. }
